Stack Safety as a Security Property

### 2020-07-10 main.tex
CP

In summary, we offer the following contributions:
\begin{itemize}
\item
  % \leo{More? Extensional}\bcp{We might need to invent a word for   this---''quasi-extensional''?}
  A new {\em formal stack-safety property} capturing the
  intuitive concepts of strict local state encapsulation and well-bracketed
  control flow~(\cref{sec:lse-and-wbcf}) in terms of familiar security concepts.
  % These properties   rule out common examples of stack memory abuse.
  This property can be enforced by
  the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, which is
  based on hardware tagging~(\cref{sec:conservative}).
\item
  A {\em lockstep} variant of the stack-safety property that is also
  enforced by the Depth Isolation policy
  but is more efficient to test, plus a proof (in Coq) that this variant is
  stronger than the original~(\cref{sec:lockstep}).
\item
  A {\em lazy} variant of the stack-safety property that can be enforced by the
  more efficient Lazy Tagging and Clearing policy
  of \citet{DBLP:conf/sp/RoesslerD18}, plus a proof (in Coq) that this
  variant is weaker than the original~(\cref{sec:lazy}).
%  This policy does not enforce our strict
%  local state encapsulation property,
%  but a modified version of it using unique per-activation tags enforces
%  an (also provably) weaker lazy variant of the property, which still prohibits
%  the same intuitively stack unsafe behaviors.
%
% \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
%   but the point we can make is still strong: To make the property easier
%   to enforce, we can slightly weaken its guarantees; the weaker version is
%   a bit more complicated to state but still retains the pragmatic
%%   protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
%   depth-based tags, which do not enforce even our lazy policy. So we need to
%   introduce a version of LT\&C that uses unique activation tags.}

@@

We assume we are given a machine language program, together with certain auxiliary declarations
about its structure: the code footprint of each function and the locations of instructions
representing calls, returns, and function entry points.
Although this information would typically come from a compiler, we do not
assume anything about the provenance of the machine
code. We assume that the machine has a program counter $\PCname$
and stack pointer $\rsp$ and can perform the usual
loads, stores, and (direct or indirect) jumps; other details of the machine's ISA are unimportant.

To make our stack safety definitions precise, we make some simple assumptions about
memory layout; our definitions could easily be changed to be parametric over
these assumptions, at the cost of greater complexity.  Memory
consists of an instruction region and a stack that grows upwards; for simplicity,
we omit a heap. (Integrating a memory safety property for the heap in the style of
\citet{DBLP:conf/post/AmorimHP18} is a natural extension that we discuss
in~\Cref{sec:relwork}.)  The stack has a conventional structure, with
contiguously allocated per-function frames.
The boundary between the frames of a caller and callee is given by the $\rsp$ register
at the time of the call.  The stack can be used to
pass arguments or return results in the usual way, by placing them
at the top of the caller's frame. Functions can also communicate via registers; indeed, our
definitions say nothing at all about registers (except $\PCname$ and $\rsp$).

Our stack safety concepts depend critically on the notions of ``call'' and ``return,''
which are not fully explicit in machine code.  We define a call to be
any instruction declared as such in the program.
If a call occurs at $\PCname = a$ and $\rsp = s$, then the \emph{corresponding
  return} has occurred when $\PCname = \nextPC(a)$ and $\rsp = s$. (Our definition
of well-bracketed control flow relates this definition of return to the program's
declared set of return points.)
Our property definitions make no other detailed assumptions about call or return sequences, although
some of the enforcement mechanisms we consider do.

>>>>

In summary, we offer the following contributions:
\begin{itemize}
\item
  % \leo{More? Extensional}\bcp{We might need to invent a word for   this---''quasi-extensional''?}
  A new {\em formal stack-safety property} capturing the
  intuitive concepts of strict local state encapsulation and well-bracketed
  control flow~(\cref{sec:lse-and-wbcf}) in terms of familiar security concepts.
  % These properties   rule out common examples of stack memory abuse.
  This property can be enforced by
  the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, which is
  based on hardware tagging~(\cref{sec:conservative}).
\item
  A {\em lockstep} variant of the stack-safety property that is also
  enforced by the Depth Isolation policy
  but is more efficient to test, plus a proof (in Coq) that this variant is
  stronger than the original~(\cref{sec:lockstep}).
\item
  A {\em lazy} variant of the stack-safety property that can be enforced by
  (a variant of\bcp{same concern}) the
  more efficient Lazy Tagging and Clearing policy
  of \citet{DBLP:conf/sp/RoesslerD18}, plus a proof (in Coq) that this
  variant is weaker than the original~(\cref{sec:lazy}).
%  This policy does not enforce our strict
%  local state encapsulation property,
%  but a modified version of it using unique per-activation tags enforces
%  an (also provably) weaker lazy variant of the property, which still prohibits
%  the same intuitively stack unsafe behaviors.
%
% \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
%   but the point we can make is still strong: To make the property easier
%   to enforce, we can slightly weaken its guarantees; the weaker version is
%   a bit more complicated to state but still retains the pragmatic
%%   protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
%   depth-based tags, which do not enforce even our lazy policy. So we need to
%   introduce a version of LT\&C that uses unique activation tags.}

@@

We assume we are given a machine language program, together with some auxiliary information
about its structure: the code footprint of each function and the locations of instructions
representing calls, returns, and function entry points.
Although this information would typically come from a compiler together with
the machine code, we do not
assume anything about their provenance.  We assume that the machine has a
program counter $\PCname$ 
and stack pointer $\rsp$\bcp{but not a frame pointer?!}; other details of the machine's ISA are unimportant.

To make our stack safety definitions precise, we make some straightforward
assumptions about 
memory layout; our definitions could easily be changed to be parametric over
these assumptions, at the cost of some complexity in their statements.  Memory
consists of an instruction region and a stack that grows upwards. (For simplicity,
we do not consider the heap. Integrating a memory safety property for the
heap in the style of 
\citet{DBLP:conf/post/AmorimHP18} is a natural extension that we discuss
in~\Cref{sec:relwork}.)  The stack has a conventional structure, with
contiguously allocated per-function frames.
The boundary between the frames of a caller and callee is indicated by the $\rsp$ register
at the time of the call.  The stack can be used to
pass arguments or return results in the usual way, by placing them
at the top of the caller's frame. Callers and callees can also communicate
via registers if they wish; indeed, our 
definitions say nothing at all about registers (except $\PCname$ and
$\rsp$). \bcp{This implicitly amounts to choosing a caller-saves register
  convention, though, doesn't it?}

Our stack safety concepts depend critically on the notions of ``call'' and ``return,''
which are not fully explicit in machine code (the same instruction may be
used in some places as part of a call or return sequence and in other places
for other purposes).  We define a call to be
any instruction declared as such in the program\bcp{in the call map, not the
program, right?}.
If a call occurs at $\PCname = a$ and $\rsp = s$, then the \emph{corresponding
  return} has occurred when $\PCname = \nextPC(a)$ and $\rsp = s$. (Our definition
of well-bracketed control flow relates this definition of return to the program's
declared set of return points.)
Our property definitions make no other detailed assumptions about call or return sequences, although
some of the enforcement mechanisms we consider do.
